Child survival for mothers: mortality change and related measures

### 2019-08-13 draft.Rmd
Cambios luego de sugerencias de Diego. Incorporé gráfico de cambio relativo y extendí un poco la explicación en las medidas relativas.

---
title: "Changes in mortality and child survival for mothers"
author: "Ivan Williams and Diego Alburez-Gutierrez"
date: "07 August 2019"
output: html_document
bibliography: bib.bibtex
---

@@

A discrete approximate of the effect on $CS_a$ given an absolute change in mortality $m_{x,\delta}=m_{x}+\delta$ is:

@@

The effect of a relative change $\delta$ will be bigger when the mortality is more disperse in ages where the fertility is very concentrated. For example in \ref{fig:relative_change}, the youngest schedule would have a bigger change in $C_{CS_a}$...

```{r relative_change, echo=F, warning=F, fig.cap="Two different fertility schedules (black lines) and a big hump mortality case (right axis, line in red). In dashed grey lines are the mean age at birth"}

@@

Considering that $m_{x,\delta}=m_{x}+\delta$ ((@wrycza2012)), and $l_{a-x} = e^{-\int_{0}^{a-x}{(\mu_t+\delta})}$ :

@@

@@

This section proposes three additional measures related to the child survival measure for a mother aged $a$.

*Mean Time Spent with Lost* ($MTSL$) is an absolute measure of the expected time with a death daughter that lives a mother, which can be expressed in terms of a temporary expected lost years measure $e^\dagger$:

@@

It is reasonable to think that mothers relates these lost years with the expected time with live daughters, getting the *intensity* of her losts (*Intensity Time Lost* ($ITL$)). This makes possible also to compare between fertility regimes:

@@

Different populations could have different mean age at lost ($MAL$) of mothers with death child:

@@

$K_a$ is the mean age at birth for womens aged $a$, $MAD_{a-x}$ refers to the mean age at death for those newborns who dies before $a-x$, $TFR_a$ is the cumulated fertility for a women aged a, and $f_{x,a}$ is the cumulative distribution of fertility until age $a$.

Equation \ref{eq:MAL} can be derived by starting with the integral of the age $x+t$ at each death child weighted by fertility and survival function:

\begin{equation}
\label{eq:MALproof}
\begin{split}  
  MAL_a = \frac{\int_\alpha^a{m_x {\int_0^{a-x}\frac{d_t (x+t) dt}{\int_0^{a-x}{d_tdt}}}}}{\int_\alpha^a{m_x}} \\
  MAL_a = \frac{\int_\alpha^a{m_x x}}{TFR_a} + \frac{\int_\alpha^a{m_x MAD_{a-x}}}{TFR_a} \\
  MAL_a = K_a + \frac{\int_\alpha^a{m_x MAD_{a-x}}}{TFR_a} \\
  MAL_a = K_a + \int_\alpha^a{f_{x,a} MAD_{a-x}}
\end{split}
\end{equation}

>>>>

---
title: "Child survival for mothers: mortality change and related measures"
author: "Ivan Williams and Diego Alburez-Gutierrez"
date: "07 August 2019"
output: pdf_document
bibliography: bib.bibtex
---

@@

A discrete approximate of the effect on $CS_a$ given an absolute change in mortality $m_{x,\delta}=m_{x}+\delta$ is (@wrycza2012):

@@

The effect of a relative change $\delta$ will be bigger when the mortality is more disperse in ages where the fertility is more concentrated. For example in \ref{fig:relative_change}, with a big hump in first adult ages, the youngest schedule would have a bigger change in than the older one in $C_{CS_a}$.

```{r relative_change, echo=F, fig.align='center', warning=F, out.width = '70%', fig.cap="Two different fertility schedules (black lines) and a big hump mortality case (right axis, line in red). In dashed grey lines are the mean age at birth for each pattern."}

@@

Considering that $m_{x,\delta}=m_{x}+\delta$ (@wrycza2012), and $l_{a-x} = e^{-\int_{0}^{a-x}{(\mu_t+\delta})}$ :

@@

This section proposes two additional measures related to the child survival measure for a mother aged $a$.

@@

*Mean Time Spent with Lost* ($MTSL$) is an absolute measure of the expected total time with a death daughter that lives a mother, which can be expressed in terms of a temporary expected lost years measure $e^\dagger$:

@@

It is reasonable to think that mothers relates these lost years with the expected time with live daughters, getting the *intensity* of her losts (*Intensity Time Lost* ($ITL$)): ratio between expected time with "lost"" and expected time with "life". This makes possible also to compare between fertility regimes:

@@

Different populations could have different mean age at lost ($MAL_a$) for mothers aged $a$ with at least a death child:

@@

$K_a$ is the mean age at birth for womens aged $a$, $MAD_{a-x}$ refers to the mean age at death for those newborns who dies before $a-x$, $TFR_a$ is the cumulated fertility for a women aged a, and $f_{x,a}$ is the cumulative age distribution of fertility until age $a$.

Equation \ref{eq:MAL} can be derived by starting with the mother age $x+t$ at each death child age at death $t$, weighted by the fertility and survival function:

\begin{equation}
\label{eq:MALproof}
\begin{split}  
  MAL_a = \frac{\int_\alpha^a{m_x {\int_0^{a-x}\frac{d_t (x+t) dt}{\int_0^{a-x}{d_tdt}}}}}{\int_\alpha^a{m_x}} \\
  MAL_a = \frac{\int_\alpha^a{m_x x}}{TFR_a} + \frac{\int_\alpha^a{m_x MAD_{a-x}}}{TFR_a} \\
  MAL_a = K_a + \frac{\int_\alpha^a{m_x MAD_{a-x}}}{TFR_a}
\end{split}
\end{equation}

Arriving to the equation \ref{eq:MAL}.
